Parsing Postfix log files

### 2007-07-02 doc/logparser.tex
Improved the explanation of dips in the normal/shuffle/reverse factor graph.

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
for this took some time to arrive at, but it turns out to be reasonably
simple.  Examining the database after parsing logfile~22 shows 64 rules
matched during parsing, but the distribution of matching rules is extremely
skewed, as shown in graph~\ref{log22 distribution by rule}.  Comparing the
number of log lines per program versus rules per program in
graph~\ref{log22 line and rule distribution by program} shows that
\textit{postfix/qmgr\/} has the highest number of log lines; however it has
a low number of rules, so the average parsing time for a
\textit{postfix/qmgr\/} log line is quite low, reducing the overall average
parsing time per log line.

XXX FIGURE OUT THE DIP AT LOG 34, mail.log-20070228

@@

Two large dips can be seen in graph~\ref{normal regex vs disarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graph~\ref{execution time vs file size vs number lines}.  The
anomaly at files 62--68 is caused by a combination of a mail loop and a
delivery failure which resulted in a huge number of log lines, all of which
were matched by a single rule.  Sorting the rules by the number of
successful matches has the effect that this rule will be tried first, so
the average number of regexs compiled per line drops sharply for those
seven log files, resulting in a much lower increase in relative execution
time, as detailed below.

>>>>

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
for this took some time to arrive at, but it turns out to be reasonably
simple.  Examining the database after parsing logfile~22 in isolation shows
64 rules matched during parsing, but as shown in graph~\ref{log22
distribution by rule} the distribution of matching rules is extremely
skewed.  Comparing the number of log lines per program versus rules per
program in graph~\ref{log22 line and rule distribution by program} shows
that \textit{postfix/qmgr\/} has the highest number of log lines, however
it has a low number of rules.  This combination results in a low average
number of rules to try for \textit{postfix/qmgr\/} log lines, reducing the
overall average number of rules to try per log line.  In turn this reduces
the average parsing time per log line, and thus the parsing time for the
log file.  While the parsing time is substantially greater than the time
for the log files on either side, the increase from optimal ordering to
reversed ordering is proportionally smaller, producing the dips in
graph~\ref{normal vs shuffled vs reversed ordering factor}.

@@

Two large dips can be seen in graph~\ref{normal regex vs disarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graph~\ref{execution time vs file size vs number lines}.  The
anomaly at files 62--68 is caused by a combination of a mail loop and a
delivery failure which resulted in a huge number of log lines, all of which
were matched by a single rule.  Sorting the rules by the number of
successful matches has the effect that this rule will be tried first, so
the average number of regexs compiled per line drops sharply for those
eight log files, resulting in a much lower increase in relative execution
time, as detailed below.
